PAMELA: an annotation-based Java Modeling Framework

### 2021-02-02 architecture.tex
typos

To summarize, in the following we list the most common Java annotations used in the context of \texttt{ModelEntity} and \texttt{ModelProperty} definitions:

@@

PAMELA model at runtime is computed dynamically, working on the classpath of launched Java application, and starting from a simple Java interface (or a collection of Java interfaces) which is/are PAMELA-annotated. From a mathematical point of view, internal representation of the underlying model is a graph whose vertex are PAMELA \emph{ModelEntities} (annotated Java interface), and edges are either inheritance links or reference links (a property whose type is another \emph{ModelEntity}. \mytexttt{@Imports} and \mytexttt{@Import} annotations allows to include some other \emph{ModelEntities} in the model. On the contrary, an annotation attribute \mytexttt{@Getter(...ignoreType=true)} allows to ignore the link. In that context, PAMELA model computation is a graph closure computation, starting from a collection of vertices.
%\footnote{Model closure computation on-the-fly provides an interesting approach to deal with model fragmentation.} 
A PAMELA model at runtime is represented by a \mytexttt{ModelContext}.

@@

% Folloging section/itemize may be removed when needed 
%This composition offers many benefits: 
%\begin{itemize}
%    \item Strong coupling between model and code
%    \item Strong typing is kept, and required checks are performed by the Java compiler
%    \item PAMELA framework provides interpretation of model@runtime
%    \item No need to generate POJO (plain old Java objects), as their execution follow the standard semantics (less code, less bugs)
%    \item Custom implementation are provided if needed, using classical Java extension points
%    \item It offers a way to intercept method calls and instrument the code
%    \item Assertions checking at runtime
%    \item Dynamic code weaving at runtime (aspect programming without compilation)
%\end{itemize}

>>>>

In the following we list the most common Java annotations used in the context of \texttt{ModelEntity} and \texttt{ModelProperty} definitions:

@@

PAMELA model at runtime is computed dynamically, working on the classpath of
launched Java application, and starting from a simple Java interface (or a
collection of Java interfaces) which is/are PAMELA-annotated. From a
mathematical point of view, internal representation of the underlying model is
a graph whose vertices are PAMELA \emph{ModelEntities} (annotated Java
interface), and edges are either inheritance links or reference links (a
property whose type is another \emph{ModelEntity}). \mytexttt{@Imports} and
  \mytexttt{@Import} annotations allows to include some other
  \emph{ModelEntities} in the model. On the contrary, an annotation attribute
  \mytexttt{@Getter(...ignoreType=true)} allows to ignore the link. In that
  context, PAMELA model computation is a graph closure computation, starting
  from a collection of vertices.

%\footnote{Model closure computation on-the-fly provides an interesting approach to deal with model fragmentation.} 
A PAMELA model at runtime is represented by a \mytexttt{ModelContext}.

@@

% Folloging section/itemize may be removed when needed 
%This composition offers many benefits: 
%\begin{itemize}
%    \item Strong coupling between model and code
%    \item Strong typing is kept, and required checks are performed by the Java compiler
%    \item PAMELA framework provides interpretation of model@runtime
%    \item No need to generate POJO (plain old Java objects), as their
%    execution follow the standard semantics (less code, fewer bugs)
%    \item Custom implementation are provided if needed, using classical Java extension points
%    \item It offers a way to intercept method calls and instrument the code
%    \item Assertions checking at runtime
%    \item Dynamic code weaving at runtime (aspect programming without compilation)
%\end{itemize}

### 2021-02-02 related.tex
typos

Very few approaches build interfaces between MDE artifacts and source code at runtime. FAME \cite{kuhn2008fame} is  a library that keeps metamodels accessible and adaptable at runtime. Basically, FAME attaches meta-information to the objects of a running application permitting an uniform manipulation of both models and system objects at runtime. Similarly, in \cite{song2010applying} the authors use runtime models to maintain a causal connection between a model and a running systems for monitoring and control purposes. More recently, in \cite{boronat2019} the authors construct and maintain at runtime model-based views on the data manipulated in object-oriented code. Different to the aforementioned approaches, PAMELA does not focus in reflecting and manipulating running systems but in: 1) extending their behaviors; and 2) maintaining a continuous cycle of metamodeling so that models and corresponding code remain synchronized without the need for code generation processes. In a different approach, UMPLE \cite{lethbridge2016merging} mixes programming and modeling by integrating UML constructs into languages such as Java. However, they use code generation for the runtime part of the system.

>>>>

Very few approaches build interfaces between MDE artifacts and source code at
runtime. FAME \cite{kuhn2008fame} is  a library that keeps metamodels
accessible and adaptable at runtime. Basically, FAME attaches meta-information
to the objects of a running application permitting a uniform manipulation of
both models and system objects at runtime. Similarly, in
\cite{song2010applying} the authors use runtime models to maintain a causal
connection between a model and a running systems for monitoring and control
purposes. More recently, in \cite{boronat2019} the authors construct and
maintain at runtime model-based views on the data manipulated in
object-oriented code. Different to the aforementioned approaches, PAMELA does
not focus in reflecting and manipulating running systems but in: 1) extending
their behaviors; and 2) maintaining a continuous cycle of metamodeling so that
models and corresponding code remain synchronized without the need for code
generation processes. In a different approach, UMPLE
\cite{lethbridge2016merging} mixes programming and modeling by integrating UML
constructs into languages such as Java. However, they use code generation for
the runtime part of the system.
